POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types

### 2023-10-12 challenges/coinduction.tex
added some proof sketches

\begin{theorem}
  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
\end{theorem}
\begin{proof}
  % We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
  % commutative.
  We prove the three properties separately: 
  \begin{itemize}
  \item Reflexivity is straightforward: for any $P$, we need to show
    that $P\sbbisim{}{} P$. Both conditions in
    equations~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
  \item Symmetry is similar to the previous case: for both
    $P\sbbisim{}{} Q$ and $Q\sbbisim{}{} P$ the same conditions apply.
  \item For transitivity, we need to prove that if $P\sbbisim{}{} Q$
    and $Q\sbbisim{}{} R$ then $P\sbbisim{}{} R$. If
    $\observable{P}{\mu}$ then, by $P\sbbisim{}{} Q$,
    $\observable{Q}{\mu}$. Hence, by $Q\sbbisim{}{} R$,
    $\observable{R}{\mu}$. Moreover, if $\transition{P}{\Atau}{P'}$
    there exists $Q'$ such that $\transition{Q}{\Atau}{Q'}$ and
    $P'\sbbisim{}{}Q'$. Also, $\transition{R}{\Atau}{R'}$ with
    $Q'\sbbisim{}{}R'$. Coinductively, $P'\sbbisim{}{}R'$, hence
    $P\sbbisim{}{}R$.
  \end{itemize}

@@

\begin{lemma}
  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
\end{lemma}

>>>>

\begin{theorem}
  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
\end{theorem}
\begin{proof}
  % We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
  % commutative.
  We prove the three properties separately: 
  \begin{itemize}
  \item Reflexivity is straightforward: for any $P$, we need to show
    that $P\sbbisim{}{} P$. Both conditions in
    equations~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
  \item Symmetry follows immediately since bisimilarity is defined as
    a {\em symmetric} relation.
  \item For transitivity, we need to prove that if $P\sbbisim{}{} Q$
    and $Q\sbbisim{}{} R$ then $P\sbbisim{}{} R$. In order to do so,
    we prove that the relation
    $\mathcal R=\{(P,R)\mid \exists Q\text{ such that }P\sbbisim{}{} Q
    \land Q\sbbisim{}{} R\}$ is a strong barbed bisimulation. Let us
    assume that $(P,R)\in\mathcal R$. Hence, there exists a $Q$ such
    that $P\sbbisim{}{} Q$ and $Q\sbbisim{}{} R$. Clearly, if
    $\observable{P}{\mu}$ then, by $P\sbbisim{}{} Q$,
    $\observable{Q}{\mu}$. And, by $Q\sbbisim{}{} R$,
    $\observable{R}{\mu}$. Moreover, if $\transition{P}{\Atau}{P'}$
    there exists $Q'$ such that $\transition{Q}{\Atau}{Q'}$ and
    $P'\sbbisim{}{}Q'$. And also, $\transition{R}{\Atau}{R'}$ with
    $Q'\sbbisim{}{}R'$. Finally, by definition of $\mathcal R$,
    $(P',R')\in\mathcal R$.
  \end{itemize}

@@

\begin{lemma}
  \( \sbcong{}{} \) is the largest congruence included in
  \( \sbbisim{}{} \).
\end{lemma}
\begin{proof}
  By definition, \( \sbcong{}{} \) is a congruence and is included in
  \( \sbbisim{}{} \). Also, if it were strictly included in another
  relation then we would get a contradiction since it must be closed
  under all contexts.
\end{proof}
